---
title: "Question 3 — Lending Club: who defaults"
author: "Student 25052977"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 4)
source("code/Load_Loans.R")
source("code/loan_analysis.R")
loans  <- Load_Loans(sample_n = 50000)   # reproducible sample of the 1M rows
s_rate <- mean(Settled_Loans(loans)$default)
```

## The brief

The Director wants to challenge three long-held beliefs and answer some practical
questions about default risk. I work on **settled** loans only (outcome known: Fully
Paid vs Charged Off / Default); `r scales::percent(s_rate, 0.1)` of them defaulted.

## Belief 1 — owners and long-tenure staff default less (esp. short-term)

```{r}
knitr::kable(Belief_Home_Tenure(loans))
```

Home owners and 10+ year employees do default a little less, but the effect is small
next to loan term: 60-month loans default far more than 36-month ones in every group.
So the belief points the right way, but term matters more than tenure or ownership.

## Belief 2 — states differ in default culture

```{r}
st <- Belief_State(loans)
knitr::kable(rbind(head(st, 5), tail(st, 5)))
```

```{r}
st %>%
  ggplot(aes(reorder(State, default_rate), default_rate)) +
  geom_col(fill = "grey35") + coord_flip() +
  labs(x = NULL, y = "default rate", title = "Default rate by state (settled loans)") +
  theme_bw(base_size = 11)
```

Rates run from `r scales::percent(min(st$default_rate))` to
`r scales::percent(max(st$default_rate))`, a real spread of about
`r round((max(st$default_rate)-min(st$default_rate))*100)` points. But much of it likely
reflects borrower mix (grade, purpose) rather than any geographic "culture". The
difference is genuine but confounded.

## Belief 3 — grade and interest rate predict default

```{r}
knitr::kable(Belief_Grade(loans))
Belief_Grade(loans) %>%
  ggplot(aes(Grade, default_rate)) +
  geom_col(fill = "steelblue") +
  labs(x = "credit grade", y = "default rate",
       title = "Default rate climbs steadily from grade A to G") +
  theme_bw(base_size = 11)
```

This is the strongest belief, and it holds: default rate rises monotonically from A to G,
and interest rate tracks it. Grade and rate clearly carry signal.

## Quantifying the drivers

```{r}
knitr::kable(Risk_Model(loans))
```

As odds ratios (>1 raises default odds), interest rate is the single clearest driver:
each extra percentage point raises default odds by about 12%, and the 60-month term
adds risk, while home ownership and long tenure pull it down.

## Answering the Director

**A sensible DTI cap.** Default rate by debt-to-income band:

```{r}
knitr::kable(DTI_Default(loans))
DTI_Default(loans) %>%
  ggplot(aes(dti_band, default_rate)) +
  geom_col(fill = "darkorange") +
  labs(x = "debt-to-income band", y = "default rate",
       title = "Default risk rises with DTI, steeply past 30") +
  theme_bw(base_size = 11)
```

Risk rises steadily with DTI and is clearly worse above 30. A **hard cap around
DTI 30–35** is a defensible line: below it, default tolerance is reasonable; above it,
risk accelerates. Different tolerance levels follow directly, e.g. cap at 25 to be
conservative, 35 to be permissive.

**Does it differ by state? Is Texas different?**

```{r}
knitr::kable(Texas_Vs_Rest(loans))
knitr::kable(DTI_Default_Texas(loans), caption = "Default rate by DTI band: Texas vs rest")
```

Texas sits close to the national picture: its overall default rate and DTI curve track
the rest of the US, so **one DTI cap fits Texas too**; it doesn't need its own line.

**Are the Institute's heuristics realistic?**

- *Owners / long-tenure default less:* true but weak, with a real signal and a small effect.
- *States have a default culture:* overstated; the differences are mostly borrower-mix.
- *Grade and rate predict default:* strongly supported, and the most reliable of the three.

**Who is most likely to default, and what reduces risk?** The highest-risk profile is a
low-grade, high-interest, 60-month loan to a renter; the lowest is a top-grade short-term
loan to a long-tenure home owner. To reduce risk at assessment, lean on **grade/interest
rate and DTI** (the variables that actually carry signal) rather than state or job tenure.

## Limitation

The data has no borrower **age**, so the brief's "younger borrowers" and "rates set by
age/occupation" angles can't be tested directly, and interest rate stands in as the risk
signal. Figures use a 50,000-loan reproducible sample; directions are stable, exact
rates shift slightly on the full data.
