---
title: "Question 4 — Netflix: what works in streaming"
author: "Student 25052977"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 4)
source("code/Load_Netflix.R")
source("code/netflix_analysis.R")
nf <- Load_Netflix()   # ../data/netflix/titles.rds, list-columns parsed
```

## The brief

A team thinking about launching a streaming service wants a quick read on what sits on
Netflix and what rates well: the mix of content by country and type, the genres that
dominate, how titles score, how long movies run, and what the catalogue is "about".

## What's on the platform

```{r}
knitr::kable(Type_Split(nf), caption = "Movies vs shows")
```

```{r}
knitr::kable(Country_Summary(nf, 10), caption = "Top production countries by number of titles")
```

The catalogue is mostly films, and heavily US-led, with India, the UK, Japan and South
Korea making up most of the rest. For a new entrant that says the obvious markets are
crowded, and that Korean and Japanese content is a meaningful slice worth watching.

## Which genres dominate, and do they rate well?

```{r}
gs <- Genre_Summary(nf, 12)
gs %>%
  ggplot(aes(reorder(genre, titles), titles)) +
  geom_col(fill = "firebrick") +
  coord_flip() +
  labs(x = NULL, y = "number of titles",
       title = "Most common genres on Netflix") +
  theme_bw(base_size = 11)
knitr::kable(gs, caption = "Genre counts and average IMDb score")
```

Drama and comedy are by far the most stocked, but volume and quality don't line up:
the most common genres sit around the middle on IMDb, while smaller genres like
documentation and history tend to rate higher. Stocking more of a genre does not make
it better content.

## How do titles rate?

```{r}
knitr::kable(Ratings_By_Type(nf), caption = "IMDb rating by type")
nf %>%
  filter(!is.na(imdb_score)) %>%
  ggplot(aes(imdb_score, fill = type)) +
  geom_density(alpha = 0.4) +
  labs(x = "IMDb score", y = NULL, title = "Shows rate a little higher than movies") +
  theme_bw(base_size = 11)
```

Shows tend to score slightly higher than movies, and the spread is wide for both:
plenty of low-rated filler alongside the hits. A new service competing on quality could
do well to be more selective than the "stock everything" approach.

## How long are the movies?

```{r}
Movies_Runtime(nf) %>%
  ggplot(aes(runtime)) +
  geom_histogram(binwidth = 5, fill = "steelblue") +
  labs(x = "runtime (minutes)", y = "movies",
       title = "Most movies cluster around 90–100 minutes") +
  theme_bw(base_size = 11)
```

Movie length is tightly clustered around the standard 90–100 minute feature, tailing off
past two hours. There's little appetite on the platform for very short or very long films.

## What is the catalogue about?

```{r}
Description_Words(nf, 20) %>%
  ggplot(aes(reorder(word, n), n)) +
  geom_col(fill = "grey35") +
  coord_flip() +
  labs(x = NULL, y = "mentions in descriptions",
       title = "Most common words in title descriptions") +
  theme_bw(base_size = 11)
```

The descriptions lean on life, family, love, young and friends: the themes the platform
keeps coming back to. Useful for a new service deciding what emotional territory to own.

## Takeaways for the team

- The market is film-heavy and US-dominated, but Korean and Japanese content is a real,
  growing slice worth targeting.
- Stocking a genre heavily doesn't make it rate well: documentaries and history punch
  above their volume on IMDb.
- Shows edge out movies on ratings; quality varies widely, so curation beats quantity.
- Movies cluster at 90–100 minutes; descriptions centre on family, love and youth.
