---
title: "Question 1 — Coffee Hub: what to stock"
author: "Student 25052977"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Functional programming: this report only *calls* functions defined in code/.
source("code/Load_Coffee.R")
source("code/coffee_keywords.R")
source("code/coffee_value.R")
source("code/coffee_regions.R")

coffee   <- Load_Coffee()              # reads + cleans ../data/Coffee/Coffee.csv
coffee   <- Score_Keywords(coffee)     # adds kw_hits / kw_density
pq       <- Price_Quality_Cor(coffee)  # does price buy quality?
regions  <- Region_Summary(coffee)
roasters <- Roaster_Summary(coffee)
rs       <- Roast_Summary(coffee)
```

## The brief

Advise a coffee entrepreneur which **characteristics, regions and suppliers** to stock,
using `r nrow(coffee)` expert reviews matched against what locals actually like.

## Data at a glance

- `r nrow(coffee)` reviews, `r ncol(coffee)` variables.
- Survey keywords scored per review (`kw_hits`): mean
  `r round(mean(coffee$kw_hits), 2)`, max `r max(coffee$kw_hits)`.

## Does paying more buy a better cup?

- Spearman correlation between cost and rating: **rho = `r pq$rho`** (p = `r pq$p_value`).
- `r if (pq$rho < 0.15) "Essentially no link; price is a weak guide to quality." else "Some positive link, but far from deterministic."`
- Implication: stock on **value**, not on price tag.

## Best value coffees (rating per unit cost)

```{r}
knitr::kable(Top_Value(coffee, 10))
```

## Best origin regions

```{r}
knitr::kable(head(regions, 8))
```

## Best roasters / suppliers

```{r}
knitr::kable(head(roasters, 8))
```

## Which roast to stock?

```{r}
knitr::kable(rs)
```

- Most reviewed coffees are **Medium-Light** roasts, the local specialty default.
- Best-rated roast: **`r as.character(rs$Roast[which.max(rs$avg_rating)])`**
  (avg rating `r max(rs$avg_rating)`).
- Lighter roasts score best at the specialty end; stock mainly Medium-Light with a
  Light option for the top scorers, and keep dark roasts minimal.

## What locals like — recommended shortlist

```{r}
knitr::kable(Good_Local_Picks(coffee, 8))
```

## Recommendation

- **Stock for value over price:** cost only loosely predicts rating (rho = `r pq$rho`).
- **Lead with `r regions$Region[1]` and `r regions$Region[2]`:** highest average
  ratings among well-reviewed origins.
- **Favour Medium-Light / Light roasts**, which carry the local specialty market.
- **Anchor suppliers on `r roasters$Roaster[1]`** (avg rating `r roasters$avg_rating[1]`,
  `r roasters$n[1]` reviews).
- **Curate the "good local" shortlist** above: coffees that hit the survey descriptors
  locals use *and* deliver value.
